|

### 2023-08-16 chapter_4/src/prr/dcr_color_and_non-political_behaviors_prr.qmd
doing some more thinking

@@

Building on this set of evidence, a growing area of the literature suggests that we are able to make inferences about the partisanship about other individuals, but we are able to do this for groups of people through the increasing politicultural distinctions between Republicans and Democrats. @hetherington_prius_2018 demonstrates that Republicans and Democrats are not just different in their policy preferences but rather that they live in different places and have different lifestyles. Others have replicated and built upon this work by suggesting that while there are many different types of lifestyles that Republicans and Democrats can take on, there is increasingly clear distinction between Republicans and Democrat archetypes [@rogers_politicultural_2022]. These politicultural distinctions are clear enough that we train a computer to make accurate guesses (upwards of 80%) of the partisanship by using only images of the cars in a particular neighborhood.

Though not inherently or necessarily political, our communities often act in political ways. For example, some evidence suggests despite variation between counties' partisan composition and lockdowns discouraging people from leaving their homes during the COVID-19 pandemic, those living in Republican counties spent more time away from home than those living in Democratic counties[@roberts_polarized_2021]. Further, we see a place-based identity for rural individuals that predict partisanship [@cramer_politics_2016; see also @lyons_youre_2021] and characteristics like expression of anti-intellectual attitudes [@lunz_trujillo_rural_2022]. While place-based identities could be separate or informative of partisan identities, some literature argues that partisanship may actually inform which community we choose to make ours.

@@

Some work suggests, however, that these expressed preferences to live near copartisans do not carryover to migration patterns [see @iyengar_affect_2012]. In fact, there is some evidence suggesting that controlling for everything else but a political sorting, we would have less political homogeneity than we had in 2008 [@martin_does_2020]. Further, evidence by @martin_does_2020 suggests that people switch their party registration after moving as opposed to moving to be more aligned with their party registration.

The goal of the following sections is to discuss one potential mechanism that may help to tease out why there is weak observational empirical evidence for a dearth of theoretical and empirical evidence

>>>>

# Goals of chapter

- Demonstrate whether snap-judgments are used and how effective they are for groups of people
- Demonstrate how snap-judgments may be used even for non-political objects and the consequences
- Add one potential mechanism for the puzzle of partisan sorting: attitude expression suggests it happens, but behaviorally it does not happen much.
- Test the boundary conditions for color being a useful partisan cue

@@

Building on this set of evidence, a growing area of the literature suggests that we are able to make inferences about the partisanship about other individuals, but we are able to do this for groups of people through the increasing politicultural distinctions between Republicans and Democrats. @hetherington_prius_2018 demonstrates that Republicans and Democrats are not just different in their policy preferences but rather that they live in different places and have different lifestyles. Others have replicated and built upon this work by suggesting that while there are many different types of lifestyles that Republicans and Democrats can take on, there is increasingly clear distinction between Republicans and Democrat archetypes [@rogers_politicultural_2022]. These politicultural distinctions are clear enough that we train a computer to make accurate guesses (upwards of 80%) of the partisanship by using only images of the cars in a particular neighborhood [@gebru_using_2017].

Though not inherently or necessarily political, our communities often act in political ways. For example, some evidence suggests despite variation between counties' partisan composition and lockdowns discouraging people from leaving their homes during the COVID-19 pandemic, those living in Republican counties spent more time away from home than those living in Democratic counties [@roberts_polarized_2021]. Further, we see a place-based identity for rural individuals that predict partisanship [@cramer_politics_2016; see also @lyons_youre_2021] and characteristics like expression of anti-intellectual attitudes [@lunz_trujillo_rural_2022]. While place-based identities could be separate or informative of partisan identities, some literature argues that partisanship may actually inform which community we choose to make ours.

@@

Some work suggests, however, that these expressed preferences to live near copartisans do not carryover to migration patterns [see @iyengar_affect_2012]. In fact, there is some evidence suggesting that controlling for everything else but current rates of political sorting, we would have less political homogeneity than we had in 2008 [@martin_does_2020]. Further, evidence by @martin_does_2020 suggests that people switch their party registration after moving as opposed to moving to be more aligned with their party registration.

The literature on partisan sorting seems to conclude that people express a desire to sort into neighborhoods with partisans, but these attitudes do not turn into observable behaviors. The remainder of the chapter provides one explanation for this puzzle by demonstrating how the snap-judgment model may be at play. Much like the literature on the attitude expression of partisan sorting, the snap-judgment model heavily relies on the ways in which seemingly non-political information comes to be associated with one partisan group or another. The model also demonstrates how these snap-judgments are capable of having initial impacts on attitude expression but its effects weaken in light of explicit and countervailing information.

# Snap-judgments explaining the puzzle of political sorting

As discussed thus far in the chapter, non-political objects are often associated with particular partisan leanings. For example, the types of cars that we drive are often used and rather effective cues to one's partisanship [@hetherington_prius_2018;@gebru_using_2017]. In the same spirit as the rest of this book, we do not need to necessarily distinguish between whether someone drives a hybrid sedan and a Ford pickup truck. Information giving insight into one's partisanship may even be as subtle as their color choice in car.

Recall the snap-judgment model as depicted in @fig-theory. To apply it to the scenario of partisan sorting, absent other information about the neighborhood, the prevalence of particular colors of the cars in a neighborhood may be information that we can use to make guesses about the partisan composition of a neighborhood. Just as the other chapters demonstrated, when politics is salient, the associations between the colors red and blue trigger snap-judgments that inform attitude expression as well as behavioral reactions, and these effects appear to exist even once we have been given more explicit, and thus more informative, information about political persuasions.

{{< include ../../../assets/_theory.qmd >}}

# Experiments

- Make real estate flyers that are like a conjoint providing information about the neighborhood. Do something with a "favorite car in the neighborhood" and vary whether it is a red or a blue car.
